A Data-Driven Framework for Neural Field Modelling

### 2010-11-19 ltx/first_order_framework/LaTeX_Manuscript.tex
slightly editted kens bit on the freq analysis basis function bit

\documentclass[5p]{elsarticle}
% \documentclass[12pt]{iopart} 
% \usepackage{iopams}

@@

\parham{The additive form of the unscented Rauch-Tung-Striebel smoother} (URTSS)~\cite{Sarkka2010} is used for the state estimation. The URTSS incorporates an unscented Kalman filter (UKF)~\cite{Julier1997, Merwe2003} in a forward step to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward step to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. The sigma points, $\mathcal X_i$, are calculated using the unscented transform as follows:
\begin{align}\label{eq:sigmapoints1}
	\mathcal X_{0}&=\bar x \\
	\mathcal X_{i}&=\bar x+\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_i, \quad i=1, \dots, n_x \\
	\mathcal X_{i}&=\bar x-\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_{i- n_x}, \quad i= n_x+1, \dots, 2n_x 
\end{align}
where $\bar x$ represents either $\hat{\mathbf x}_t^{f}$ or $\hat{\mathbf x}_t^{b}$, $\mathbf{P}_x$ is the corresponding covariance matrix for filtering or smoothing, $\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_i$ is the $i^{th}$ column of the scaled matrix square root of $\mathbf P_x$, and $n_x$ is the dimension of the state space. The total number of sigma points is $2n_x+1$. The scaling parameter, $\lambda$, is defined as 
\begin{equation}\label{eq:sigmapoints3}
	\lambda=\alpha^2( n_x+\kappa) - n_x, 
\end{equation}
where the constant $\alpha$ determines the spread of the sigma points and $\kappa$ is an additional scaling parameter (see~\cite{Haykin2001} for more details).

@@

\ken{revised new bit}

To reduce the computational demand of the estimation procedure, the neural field cutoff frequency was further reduced to $\boldsymbol{\nu}_{c\phi} = 0.12$~Hz. Due to the slow frequency roll-off of Gaussian basis functions an oversampling parameter of $\rho_{\phi} = 1.67$ was chosen. Therefore for an equally spaced grid of basis functions over the full spatial domain being considered, a $9\times9$ grid was required to decompose the field, which resulted in $81$ states. The other parameter of the basis functions, $\sigma_{\phi}^2$, is set based on equation \ref{eq:CutoffFromBasisFuncWidth} and the chosen neural field bandwidth $\boldsymbol{\nu}_{c\phi}$, to give $\sigma_{\phi}^2=2.5$.

\ken{how's that?}

\dean{-------new bit start ------------}

Due to the limitations imposed by the computationally demanding estimation procedure we chose to use a $9\times9$ grid of Gaussian basis functions to decompose the field, which resulted in $81$ states.  Given the spatial domain, this yielded a basis function separation of $\Delta_{\phi}=2.5$~mm. The basis function width parameter was chosen such that a smooth representation of the field could be achieved with this arrangement, where $\sigma_{\phi}^2=2.5$. Given the basis function width, the decomposition is able to represent the field up to the spatial cutoff frequency of $\boldsymbol{\nu}_{c\phi} \approx 0.12$~Hz, as described by equation~\ref{eq:CutoffFromBasisFuncWidth}. The band-limiting effect is due to the spatial frequency response of the basis functions as seen in Fig.~\ref{fig:Figure5}(C). The result of the band-limiting effect in the field reconstruction can be observed when comparing Fig.~\ref{fig:Figure4}(A) to Fig.~\ref{fig:Figure4}(B). Given the spatial cutoff frequency of the basis functions and the basis function separation the oversampling parameter is $\rho_{\phi}\approx 1.67$.

\dean{-------new bit end ------------}

\dean{-------olde bit start ------------}

Using an oversampling parameter of $\rho_{\phi}=1$, the maximum distance between the field basis functions was calculated to be 2.5~mm to satisfy Shannon's sampling criterion (from equation~\ref{eq:BasisFunctionSeparation}). The distance between basis functions was set to $\Delta_{\phi}=2.5$~mm, which is sufficient to capture the significant dynamics according to these criteria.

Given this basis function separation, the basis function width parameter was set to $\sigma_{\phi}^2=2.5$. This allowed a smooth representation of the field by the basis function decomposition up to a spatial frequency $\boldsymbol{\nu}_{c\phi} \approx 0.12$~Hz, as described by equation~\ref{eq:CutoffFromBasisFuncWidth}. The band-limiting effect is due to the spatial frequency response of the basis functions as seen in Fig.~\ref{fig:Figure5}(C). This band-limiting effect of the basis function decomposition can be observed when comparing Fig.~\ref{fig:Figure4}(A) to Fig.~\ref{fig:Figure4}(B).

\dean{-------olde bit end ------------}

@@

\bibliographystyle{unsrt} 

>>>>

\documentclass[5p,authoryear]{elsarticle}
% \documentclass[12pt]{iopart} 
% \usepackage{iopams}

@@

The additive form of the unscented Rauch-Tung-Striebel smoother (URTSS)~\cite{Sarkka2010} was used for the state estimation. The URTSS incorporates an unscented Kalman filter (UKF)~\cite{Julier1997, Merwe2003} in a forward step to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward step to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. The sigma points, $\mathcal X_i$, are calculated using the unscented transform as follows:
\begin{align}\label{eq:sigmapoints1}
	\mathcal X_{0}&=\bar x \\
	\mathcal X_{i}&=\bar x+\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_i, \quad i=1, \dots, n_x \\
	\mathcal X_{i}&=\bar x-\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_{i- n_x}, \quad i= n_x+1, \dots, 2n_x 
\end{align}
where $\bar x$ represents either $\hat{\mathbf x}_t^{f}$ or $\hat{\mathbf x}_t^{b}$, $\mathbf{P}_x$ is the corresponding covariance matrix for filtering or smoothing, $\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_i$ is the $i^{th}$ column of the scaled matrix square root of $\mathbf P_x$, and $n_x$ is the dimension of the state space. The total number of sigma points is $2n_x+1$. The scaling parameter, $\lambda$, is defined as 
\begin{equation}\label{eq:sigmapoints3}
	\lambda=\alpha^2( n_x+\kappa) - n_x, 
\end{equation}
where the constant $\alpha$ determines the spread of the sigma points and $\kappa$ is an additional scaling parameter (see~\cite{Haykin2001} for more details).

@@

As described in Section~\ref{SpectralAnalysisSection}, the basis function configuration can be systematically chosen the account for the full spatial bandwidth from the observations, or alternatively, the configuration can be chosen to represent the neural field up to a specified bandwidth to limit the computational complexity of the estimation procedure. In the examples provided in this paper, the later option was chosen, where the neural field cutoff frequency was further reduced to $\boldsymbol{\nu}_{c\phi} = 0.12$~Hz. The result of the band-limiting effect in the field reconstruction can be seen by comparing Fig.~\ref{fig:Figure4}(A) to Fig.~\ref{fig:Figure4}(B). This effect is due to the spatial frequency response of the basis functions as seen in Fig.~\ref{fig:Figure5}(C). Due to the slow frequency roll-off of Gaussian basis functions an oversampling parameter of $\rho_{\phi} = 1.67$ was chosen. Given the spectral response of the basis functions, the over-sampling parameter, and the spatial domain, an equally spaced grid of $9\times9$ basis functions was used to decompose the field, which resulted in $81$ states. %The other parameter of the basis functions, $\sigma_{\phi}^2$, was set based on equation \ref{eq:CutoffFromBasisFuncWidth} and the chosen neural field bandwidth $\boldsymbol{\nu}_{c\phi}$, to give $\sigma_{\phi}^2=2.5$.

@@

\bibliographystyle{elsarticle-harv} 
